MILP for idling to minimise coulombic losses

### 2018-11-27 MILP_extension/MILP.tex
Create MILP.tex

\begin{document}
\linenumbers
\begin{frontmatter}

\title{Improving the Fidelity of Redox Flow Battery Schedule Optimisation for more Accurate Techno-economic Analyses.}

\author{D. Roberts }
\author[]{S. Brown \corref{cor_author}}
\ead{s.f.brown@sheffield.ac.uk}

\address{Department of Chemical and Biological Engineering, The University of Sheffield, Sheffield S10 2TN}

\cortext[cor_author]{Corresponding author}

\begin{abstract}
Redox Flow Battery (RFB) systems are promising technologies for the multi-hour electrical energy storage that would be necessary in a future low carbon emission  electrical grids based on wind and solar power. In order to perform techno-economic analyses (TEA) on these systems it is necessary to calculate potential revenue for realistic applications, ideally under optimal operation (scheduling). In this work a novel schedule optimisation framework is introduced in which coulombic and voltaic components are separated, and the problem variablised in terms of current-density rather than power. This framework allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints (NLP). It is demonstrated for an arbitrage application involving a vanadium RFB that the consideration of variable voltaic losses results in a significant improvement to the accuracy of revenue estimation when compared to the typical linear programming (LP) approach where a fixed efficiency is assumed. The introduced NLP formulation is more readily computable than alternative approaches reported in the literature, with minimal increase to solve time compared to the LP approach. The fidelity of the RFB representation is further increased by expressing open cell voltage as a function of state of charge (SOC). At low SOC, the current-density required for a given power is greater, and hence the ohmic losses are higher. In a third development, a linear constraint is applied to the working cell-voltage, resulting in current-density restriction at high SOC in order to avoid hydrogen and oxygen evolution. This feature would also be useful for schedule optimisation of Li-ion batteries, where voltage is an important consideration. The introduced framework may be further employed to assist in the optimal design of future RFB systems for specific applications.

\end{abstract}

\begin{keyword}
Redox Flow Battery \sep Optimisation \sep Non-linear Programming \sep Techno-Economic Analysis \sep Scheduling \sep VRFB  
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

% \linenumbers

%% main text
\section{Introduction}
\label{sec:Intro}
Worldwide, the need to reduce CO$_2$ emissions is stimulating the development of a wide range of renewable energy sources. %Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023 \sol{you need a reference for this}.

The growth of renewable energy is forecast to be highest in the electrical power sector, from 24\% in 2017 to 30\% in 2023. Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}%\sol{is this reference for all of the data before this? it would be better to have a couple more supporting this...}. 
Indeed, in a number of large European economies, such as Germany, UK, Italy and Spain 23\% to 30\% of electrical energy was obtained from non-hydro renewable power in 2017 \cite{BP2018}. In the US, Texas obtained around 17\% of electrical energy supply from wind power in 2017, whereas California obtained 9\% from wind and 10\% from solar\cite{ERCOT2017,California2018}.

To maintain grid stability in these regions, the fluctuating output from wind and solar is supplemented to match demand primarily by varying output from fossil fuel power thermal plant. In the UK, this is specifically achieved by gas turbines and to a lesser extent coal plant \cite{DraxElectricInsights}.% In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required at the supply side \cite{FES2018}.

The price of electricity at a given moment is set by the operating costs and finance requirements of the least economic generator that must be called upon to supply it \cite{Staffell2016}. As wind and solar power have zero marginal cost, the output of such installations depresses the electrical price intermittently. At the same time, the capacity factors of the dispatchable fossil plant will decrease, and hence a premium is added to the operating cost in order to cover CAPEX and other costs. Across the EU, the operating cost of fossil generation is further increased by the required purchase of credits under the Emissions Trading System \cite{Staffell2018}. The overall effect of these developments would be an increase in price variability, although this has been mitigated in some markets by capacity payments, made on the basis of power availability rather than energy sold \cite{ENGIE}.

Electrical price variability presents an opportunity for arbitrage by energy storage systems (ESS), which may charge and discharge in periods of low and high price respectively. At present however, this revenue alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}. Hence an important aspect of BESS economics is revenue stacking, where multiple revenue streams are captured by the same system. For example by providing grid reliability services as well as performing market arbitrage. In the UK this is illustrated by the fact that the majority of successful BESS applicants for enhanced frequency response (EFR) contracts also have capacity market contracts \cite{NGDeratingFactors}.

\cite{Sabihuddin2015} classify the services that BESS may provide by the continuous charge/discharge duration, from power quality and regulation ($<$ \SI{1}{\minute}), through bridging power (\SI{1}{\minute} to \SI{1}{\hour}) to energy management ($>$ \SI{1}{hour}). In order to satisfy a given duration a BESS must be specified with the corresponding energy to power ratio.

%At the shortest timescale, BESS may receive payments from the grid operator to respond to fluctuations in the AC grid frequency, as in the aforementioned EFR contracts. Such frequency regulation services attract Li-ion systems designed for power, such as the Hornsdale Power Reserve project in South Australia, which is specified for \SI{1}{\hour}\SI{17}{\minute} at its \SI{100}{\mega\watt} rated power output \cite{DOEGlobalESSDatabase}. This energy to power ratio is representative of global grid-scale Li-ion installation data as collated by \cite{Newbery2018}.

%An example of multi-hour energy management is the use of ESS to discharge into peak power demand in a process known as peak shaving. In California, where BESS are being considered as replacement for gas plant, a project may receive capacity payments in addition to the sale of energy. However, the regulator de-rates the payment according to a 4 hour target duration, hence a \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} BESS would only receive credit for  \SI{0.5}{\mega\watt} of power capacity \cite{NRELPeakingESS}.

%A similar de-rating has also been applied recently in the UK capacity market. This was done due to concerns that the average storage duration of the BESS contracted for the service was shorter than the \SI{2}{\hour} that a stress event may last for. \cite{NGDeratingFactors}.

%The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios.
Redox flow battery (RFB) systems are promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low. Systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}. In order to quantify the economic merits of RFB systems in comparison to other BESS, or indeed ESS, it is necessary to perform techno-economic analyses (TEA) for particular applications. TEA should involve optimisation of the operation of a BESS to maximise revenue (or minimise costs). This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}. The optimisation of operation requires the scheduling, that is the power output of the battery in each sub-period, that maximises benefit without violating energy conservation.

%Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem. In this method, the objective function and the constraints on the system are expressed as linear algebraic terms. On the resulting geometric surface, the optimum is guaranteed to lie on a vertex, hence a time-consuming search of the whole surface is not required.

\cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm. The objective was to maximise revenue on the wholesesale electrical market and the frequency response market, while always maintaining an energy reserve in order to meet . By repeating this process at different energy and power ratings, the authors found that the economically optimal VRFB rating for the \SI{50}{\mega\watt} wind farm was \SI{5.3}{\mega\watt} and \SI{3}{\mega\watt\hour}. This supports the point made above regarding the poor economics of wholesale market arbitrage - the optimal RFB rating is dictated by the frequency market revenue and the reserve energy requirement. Increasing the capacity of the VRFB in order to perform additional wholesale market arbitrage is not economical. The authors also suggested that a hybrid system may be suitable, as the economically optimal VRFB operates in a narrow SOC window when providing frequency regulation, with the remaining SOC held in reserve for large frequency deviations that occur infrequently. \cite{Vaca2017} compared a hybrid system consisting of a VRFB and a supercapacitor with the individual systems. They found that the hybrid system was the optimal choice, with the VRFB providing the energy capacity and the supercapacitor the low cost power capacity.

It is worth noting that both \cite{Johnston2015} and \cite{Vaca2017} assume a higher cost of energy capacity than power capacity for the VRFB. This contradicts economic modelling of system costs by \cite{Viswanathan2014,Ha2015Estimatingstorage}, where the energy cost is markedly lower.

%\cite{Johnston2015} and \cite{Vaca2017} apply a cost of degradation in \pound per unit energy throughput as a coefficient to ESS power variables in the objective function. This approach captures an important consideration in electrical price arbitrage using ESS - that there is a threshhold price differential below which the degradation cost outweighs the revenue. In both works the capital cost of the ESS and the degradation cost are included, but the residual value at project end is not counted. \cite{Dufo-Lopez2009} present a more detailed economic model for various ESS used to increase the value of wind power sold into a day-ahead market. The residual value of the ESS at the end of the project is pro-rated according to the number of cycles performed as a fraction of the expected cycle life, and discounted to a net present value. The schedule optimisation in this work is not as thorough as the LP approaches of  \cite{Johnston2015,Vaca2017}; the day is divided into two peak periods where the ESS is discharged and two off-peak where it is charged, and the boundaries of the periods may be varied iteratively to find an apparent optimum.

A full system LP schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a micro-grid with a load, PV, wind and dispatchable micro-generation. When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output. The authors used a mixed-integer LP (MILP) approach, where the integer variables were required to factor start-up costs. Additional global constraints were required to guarantee reliability of the micro-grid.

\cite{Gomes2017}  studied the application of a VRFB in order to cope with stochastic deviations from the forecast exports of wind and solar power into a day ahead market. A two stage optimisation was applied, and LP was used to optimise the VRFB schedule based on the forecast  in the first stage.

\cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market. This approach was not strictly LP, but the problem may be reposed in LP form.

A significant simplification in the representation of the RFB in all of the above studies is the assumption of a constant efficiency. In reality, the efficiency of any BESS is a function of the charge/discharge power, among other variables, such as state of charge (SOC). In the LP formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the  discharge SOC increment expression. This would complicate the problem as the constraint is now non-linear and non-quadratic. \cite{Nguyen2014} report a highly detailed non-linear model for the charge and discharge efficiency of a vanadium RFB. In subsequent work \citep{Nguyen2015} this function was incorporated into the SOC expression of a schedule optimisation problem in the aforementioned fashion. Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming (DP) to break the problem down. This is more computationally intensive than the LP method, but the authors state that an optimum schedule may be found in an acceptable time-scale as the small number of variables and the presence of constraints limit the search space.
A DP search approach is also used by \cite{Oudalov2007}, where the internal resistance of the BESS is expressed as a function of SOC in a TEA of a VRFB and a lead acid battery for industrial peak-shaving.

\cite{Sarker2017} also incorporated the dependency of efficiency on power in a scheduling problem (although applied to a Li-ion battery rather than a RFB). Quadratic terms are employed to relate losses to C-rate of charge and discharge. The presence of this quadratic function on the denominator of the discharge SOC increment term is avoided by redefining the discharge power variable as upstream of losses (rather than the power delivered to the grid). The discharge efficiency function is then moved from the SOC expression and multiplied by the discharge power variable in the objective function. The charge efficiency is left in the SOC expression (redefining the charge power variable in the same way would require dividing it by the quadratic charge efficiency term in the objective function). The problem therefore becomes convex (quadratic), but non-linear in both objective function and constraint. The authors then employ a linear approximation technique and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).

Accurate treatment of efficiency is important, as losses impact the revenue available from arbitrage. The key innovation of the present work is the introduction of a RFB schedule optimisation framework posed in terms of current-density rather than power. This allows the separation of the voltaic and coulombic components of the power input/output. Using the introduced framework, three improvements to the RFB representation in the schedule optimisation are posed while maintaining an easily computable problem:

\begin{itemize}
    \item Quadratic expression of ohmic losses ($I^2R$) in the objective function under linear constraints.
    \item Open cell voltage (OCV) as a function of SOC
    \item Maximum cell voltage constraint in terms of OCV and overpotential
\end{itemize}

The formulation of each of these elements and the corresponding treatment of the VFB are described in \cref{sec:Model_Formulation}. The impact of the improvements to the optimisation formulation are demonstrated in \cref{Case Study}, using the case study of a \SI{4}{\hour} rated VRFB performing arbitrage against the day-ahead N2EX electricity price.

\section{Model formulation}
\label{sec:Model_Formulation}

Pure price arbitrage is chosen as the application for schedule optimisation due to its simplicity, allowing a focus on the treatment of the RFB within the problem. In the following section, an LP formulation for schedule optimisation for maximum arbitrage revenue is first posed in current-density terms to serve as a reference. Then the three improvements to the model described in \cref{sec:Intro} are introduced. Finally, the process for specifying the RFB within the problem is defined.

\subsection{LP schedule Optimisation in Current-Density Terms}
\label{Model_Formulation_LP_Reference_Current_Density_Terms}

In the literature on RFB schedule optimisation, scheduling problems are posed using either power (\cite{Chen2012,Hu2010,Nguyen2015}) or energy (\cite{Johnston2015,Vaca2017}) as the time-indexed variables. This is not an important difference, as the former only requires that the timestep $\tau$ be included as a coefficient. Here, the power convention is adopted, as it allows the timestep to be varied independently. The objective function for pure arbitrage revenue described by \cite{Hu2010} is adopted here, but adapted so that separate charge and discharge power variables are explicitly defined. Specifying a single power variable with negative values representing charge and positive discharge is not practical for LP optimisation, as the solver will drive the variable to extremes which is likely to be sub-optimal.

In the period $Y$, made up of sub-periods $t$, the schedule is optimised by maximising the revenue defined by:

\begin{equation}
\label{eqn:Lit_Review_LP_Obj_Func}
R = \tau \sum_{t \in Y} (P_{Dis,t} - P_{Chg,t}) p_t
\end{equation}

where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power (\si{\watt}) between the RFB and the rest of the system in sub-period $t$. $p_t$ is the price of electrical energy (\SI[sticky-per, bracket-unit-denominator = false]{}[\pounds]{\per\mega\watt\hour}) in sub-period $t$, and $\tau$ the time-step parameter  (\si{\hour}).

The SOC of the RFB at time $t$ is defined after \cite{Gomes2017} by:

\begin{equation}
\label{eqn: Lit_Review_LP_SOC_Constraint_Example}
\begin{gathered}
SOC_t = SOC_{t-1} + \frac{\tau P_{Chg,t} \eta_{Chg}}{E_{BESS}} - \frac{\tau P_{Dis,t}}{E_{BESS}\eta_{Dis}} \\
\end{gathered}
\end{equation}

where $E_{BESS}$ is the parameter energy capacity and $\eta_{Chg}$ and $\eta_{Dis}$ are the charge and discharge energy efficiency parameters respectively. The SOC must at all times be constrained between a maximum of 1 and a minimum of 0, though a narrower window may be used \cite{Vaca2017}.

The above problem was reposed in terms of current-density by replacing the charge/discharge power variables with charge/discharge current-density variables, and multiplying the overall expression by a representative voltage in order to obtain power input/output by the classical $P = IV$ expression. Current-density was chosen over current as the former is a scale-independent metric commonly used in the RFB literature. The use of current-density makes multiplication by stack area necessary to obtain absolute current. Separate efficiencies representing voltaic and ohmic losses must now be specified, rather than the single energy efficiency term in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example}. As current-density is defined based on the current at the RFB terminals, coulombic losses are dealt with upstream in \cref{eqn: Method_SOC_Tracker}. It is however necessary to incorporate voltaic losses in the objective function, as these affect the power input/output.  In current-density terms, the LP objective function for arbitrage revenue (to be maximised) is defined by:

\begin{equation}
\label{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}
R_{LP} = \frac{A.\tau OCV_{50\%}}{1\times 10^{6}}\sum_{t \in Y}p_{t}(I_{D, t}\sqrt{\bar\eta_{V}}(1 - l_{BOP}) - \frac{I_{C,t}}{\sqrt{\bar\eta_{V}}(1 - l_{BOP})})
\end{equation}

where $I_{C,t}$ and $I_{D,t}$ are independent variables representing charge and discharge current density (\si{\ampere\per\square\meter}) respectively. $A$ is the stack area parameter (\si{\square\meter}) and $OCV_{50\%}$ the open cell voltage at 50\% SOC. $\bar\eta_{V}$ is the representative round-trip voltaic efficiency. The  placement of the additional fractional balance of plant losses parameter $l_{BOP}$ reflects the assumption that balance of plant power consumption is primarily due to electrolyte pumping, and that pumping power is proportional to current. The validity of this assumption is discussed in \cref{analysis_of_assumptions}.

The function describing SOC given in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example} is now expressable in purely coulombic terms. This is done by replacing the power variables with current-density variables, the energy efficiency parameters with a coulombic efficiency parameter, and the energy capacity of the RFB with a coulombic capacity. The resultant function for the SOC at the end of sub-period $t$ is defined by:

\begin{equation}
\label{eqn: Method_SOC_Tracker}
SOC_t = SOC_{t-1} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}}) \; \forall t \in Y
\end{equation}

where $C$ is the coulombic capacity of the RFB (\si{\ampere\hour}). It is assumed that the coulombic efficiency $\eta_C$ does not vary with SOC or current-density, and that coulombic losses for charge and discharge are symmetrical. This assumption is discussed in \cref{analysis_of_assumptions}.

The following constraints are applied to the list of SOC values generated by the above expression. Firstly the constraint on SOC range is formalised by:

\begin{equation}
\label{eqn: Method_Constraint_SOC_Range}
SOC_{min} \leq SOC_t \leq SOC_{max} \; \forall t \in Y
\end{equation}

where $SOC_{min}$ and $SOC_{max}$ are the minimum and maximum permitted SOC parameters.

So that the optimisation may be legitimately performed on consecutive periods of historical data, it is necessary that the SOC at the end of the period be returned to the starting value. \cite{Hu2010} set this value at 0, but in order to avoid imposing low SOC behaviour, it is here set at 0.5. This constraint is formalised by:

\begin{equation}
\label{eqn: Method_Constraint_E_Conservation}
SOC_0 = SOC_n = 0.5
\end{equation}
where $n$ is the final sub-period in $Y$.

Finally, the current-density in and out of the RFB is constrained by:

\begin{equation}
\label{eqn: Method_Constraint_Ia}
I_{min} \leq I_{D,t}, \; I_{C,t} \leq I_{max} \; \forall t \in Y
\end{equation}

where $I_{min}$ and $I_{max}$ are the minimum and maximum permitted current-densities respectively.

\subsection{NLP treatment of voltaic losses}
\label{Model_Formulation_NLP_treatment_voltaic_losses}

A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter. This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are accounted for for. In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density. The faradaic, or activation over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density \cite{Aaron2011}. \Cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation. This results in a non-linear (NLP) objective function for revenue in period $Y$ given by:

\begin{equation}
\label{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}
R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(I_{D, t}(OCV_{50\%} - V_a)(1-l_{BOP}) - \frac{I_{C, t}(OCV_{50\%} + V_a)}{1-l_{BOP}} - (I_{D,t}^{2} + I_{C,t}^{2})ASR) \; \forall t \in Y
\end{equation}

where $V_a$ and $ASR$ are scalar parameters representing faradaic over-potential (\si{\volt}) and area specific resistance (\si{\ohm\square\meter}). The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge, which is discussed further in \cref{analysis_of_assumptions}. The objective function is subject to the same constraints as described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.

Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex, as it requires maximising a negative quadratic, hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.

Under the constant coulombic efficiency assumption \cref{eqn: Method_SOC_Tracker} is linear, and so by extension are the constraints in \cref{eqn: Method_Constraint_SOC_Range,eqn: Method_Constraint_E_Conservation}.  The overall formulation hence comprises a non-linear objective function subject to linear constraints. This is an improvement over the formulations of losses expressed in power efficiency terms by \cite{Nguyen2015,Sarker2017}, both of which use non-linear SOC constraint expressions, as it is likely to be easier to resolve \cite{Hart2017}.

The formulation also benefits from a consistent definition of the current-density variables in terms of flow between the terminals and the external system, unlike in \cite{Sarker2017}.

The NLP problem is based on an assumption of constant coulombic efficiency, which is discussed further in \cref{analysis_of_assumptions}.

\subsection{Functional Treatment of OCV}
\label{Model_Formulation_Functional_Treatment_OCV}
A simplification in the NLP problem introduced in \cref{Model_Formulation_NLP_treatment_voltaic_losses} is the use of a fixed OCV, ($OCV_{50\%}$) where OCV is in fact a function of SOC. If the OCV is linear with respect to SOC in the permitted SOC range, which is the case of the system described by \cite{Reed2016} it is, and the RFB transfers as much energy below 50\% SOC as above it, then this assumption may be reasonable. However, using the current-density formulation, it is possible to functionalise the relationship between OCV and SOC. The relationship between the average OCV and the average SOC in sub-period $t$ is approximated by:

\begin{equation}
\label{eqn: OCV_vs_SOC}
OCV_t = a\frac{SOC_t + SOC_{t-1}}{2} + b 
\end{equation}

where $a$ and $b$ are system specific parameters.

Replacing $OCV_{50\%}$ with $OCV_t$ in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} and expanding $SOC_t$ and $SOC_{t-1}$ by \cref{eqn: Method_SOC_Tracker} results in an objective function  defined by:

\begin{equation}
    \label{eqn: NLP_true_OCV_long_form}
    \begin{split}
\max R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(\\
(1-l_{BOP})I_{D, t}(a\frac{SOC_{t-1} + SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} +  I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b - V_a)\\ - \frac{1}{1-l_{BOP}}(I_{C, t}(a\frac{SOC_{t-1} +  SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} + I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b + V_a))\\ 
- (I_{D,t}^{2} + I_{C,t}^{2})ASR) \; \forall t \in Y
\end{split}
\end{equation}

This objective function is non-convex, due to the presence of bivariate quadratic terms such as $-I_{D,t}I_{C,t-1}$.

\subsection{Cell Voltage Constraint}
\label{Model_Formulation_Constraint_Cell_Voltage}
It has so far been assumed that the RFB may accept the maximum charge current-density right up to the maximum permitted SOC. In reality this may result in the cell-voltage being too-high, discussed further in \cref{Results_Cell_Voltage_Constraint}. An alternative charging approach, which is common for Li-ion batteries where high cell voltages pose safety concerns, is to switch to a constant voltage taper charge at high SOC, where the current declines until a cut-off value is reached. In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical, as demonstrated by the low efficiency at low current-density reported by \cite{Nguyen2014}. Rather than globally adjust the maximum permitted charge current-density or SOC, which may lead to underestimation of revenue, the objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} is made subject to a new linear constraint in terms of cell-voltage, defined by:

    \begin{equation}
\label{eqn: Cell_Voltage_Constraint}
OCV_t + V_a + I_{C,t}ASR \leq V_{max} \; \forall t \in Y
\end{equation}

Where $V_{max}$ is the maximum permitted cell-voltage. As $OCV_t$ is an averaged value of SOC in the sub-period $t$ (\cref{eqn: OCV_vs_SOC}), the accuracy of the model should improve as $\tau$ is reduced. A similar constraint could be implemented for discharge if required.

\subsection{RFB Specification}
\label{VRFB_specification}
The RFB is modelled here as a single cell. As the power output of a hypothetical RFB may be increased at the cost of efficiency, it is necessary to specify a rated efficiency when determining the stack area required for, say, a \SI{1}{\kilo\watt} system. This is formalised by:

\begin{equation}
\label{eqn: Rated_efficiency}
\eta_{DC} = \eta_{V,rated}\eta_{C}(1-l_{BOP})^2 = \eta_{rated}
\end{equation}

Where $\eta_V$ and $\eta_C$ are the round trip voltaic and coulombic efficiencies respectively. It is assumed that the one-way fractional balance of plant losses ($l_{BOP}$) are primarily due to electrolyte pumping, and that the pump runs on DC power.

The stack area (\si{\square\meter}) is calculated by:

\begin{equation}
\label{eqn: Simple_RFB_Cost_Model_Stack_Area}
A = \frac{P_{rated}}{I_{rated}V_0\sqrt{\eta_V}(1 - l_{BOP})}
\end{equation}

where $P_{rated}$ is the system power rating (\si{\watt}) and $I_{rated}$ is the current density (\si{\ampere\per\square\metre}) corresponding experimentally to coulombic and voltaic efficiencies that satisfy \Cref{eqn: Rated_efficiency}.

The coulombic capacity of the VRFB, required for the tracking of SOC in \Cref{eqn: Method_SOC_Tracker}, is calculated by:

\begin{equation}
\label{eqn: coulombic_capacity}
C = \frac{AI_{rated}r_{e:p}}{\sqrt{\eta_C}}
\end{equation}

Where $r_{e:p}$ is the specified energy to power ratio of the system. The presence of $\eta_C$ ensures there is sufficient chemical energy in the tank to cover coulombic losses upon discharge (square root after \cite{Darling2014}).

\section{Case Study}
\label{Case Study}

\subsection{The Application}
\label{Results_The_Application}
Pure price arbitrage on the day ahead N2EX electrical market was chosen as a case study, as the availability of future price information makes deterministic schedule optimisation relevant as described by \cite{Hu2010}. Schedule optimisation was performed on 24h periods, with $\tau$ of 1h. The indexed price $p_t$ was parametrised using data obtained from the NordPool website \cite{NordPool2016}.

\subsubsection{The RFB}
\label{Model_Formulation_VRFB_Params}
\cite{Reed2016} studied various electrode flow architectures on a \SI{1}{\kilo\watt} VRFB system based on a \SI{50}{\micro\meter} Nafion NR212 membrane. In this case study, performance data were taken from the test of the highest performing stack architecture (IDD2s) at an electrolyte flow rate of 400 \si{\ml\per\minute} per cell. This setup was tested at three current-densities: 1600, 2400 and 3200 \si{\ampere\per\square\meter}. At each test condition efficiency data were obtained as integrals over a cycle between 15\% and 85\% SOC.

In the parametrisation of the VRFB for following case studies $P_{rated}$ was set at \SI{1}{\kilo\watt}, the energy to power capacity ratio $r_{e:p}$ was set at 4 and $\eta_{rated}$ was set at 75\%. $l_{BOP}$ was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}. A coulombic efficiency ($\eta_C$) of 0.975 was derived as the average of the values reported by \cite{Reed2016} at the three current-densities tested, which only varied between 0.974 and 0.976.  For the chosen system the overall energy efficiency requirement in \cref{eqn: Rated_efficiency} is satisfied when the voltaic efficiency $\eta_{V,rated}$ is 0.801, which corresponds (by linear interpolation) to a current-density ($I_{rated}$) of \SI{2188}{\ampere\per\square\meter}. $OCV_{50\%}$ was taken from \cite{Kim2011}, where the OCV of a chloride electrolyte was reported at 20\%, 50\% and 80\% SOC. Although a different electrolyte, based on a mixture of sulphuric acid and hydrochloric acid, was used in \cite{Reed2016} the OCV data have not been published.

\cref{tab:VRFB_Sizing_Params} summarises the parameter values used to represent the VRFB.

\begin{table}[!pht]
\captionsetup{font=normalsize}
\centering
\caption{Parameters describing VRFB in sizing process}
\label{tab:VRFB_Sizing_Params}
\begin{tabular}{ccccccc}
\hline
Param &  Value & Units &  Source \\
\hline
$P_{rated}$ & 1 & \si{\kilo\watt} & Arb.\\
$r_{e:p}$ & 4 & - & \multirow{2}{*}{Typical}\\
$\eta_{rated}$ & 0.75 & - \\
\hline
$\eta_{V,rated}$ & 0.801 & - &\multirow{3}{*}{\cite{Reed2016}}\\
$\eta_C$ & 0.975 & -\\
$I_{rated}$ & 2188 & \si{\ampere\per\square\meter} \\
\hline
$OCV_{50\%}$ & 1.46 & \si{\volt} & \cite{Kim2011}\\
$l_{BOP}$ & 0.02 & - & \cite{Kim2013}\\

\hline
\end{tabular}
\end{table}

\subsection{Implementation}
\label{Results_Implementation}
Each optimisation problem was posed using PYOMO within a Python loop for repeat solutions on 24h sets of data. The gurobi solver was called via the SolverFactory module. The script was run on a virtual machine (\cite{VirtualBox}) running Linux Mint, which was allocated 1024 MB RAM.

\subsection{NLP treatment of Voltaic Losses}
\label{Results_NLP_Treatment_Voltaic_losses}
In the first case study, the performance of the NLP formulation with functionalised voltaic losses (\cref{Model_Formulation_NLP_treatment_voltaic_losses}) was assessed by comparison to the LP formulation described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}. The LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} requires the assumption of a constant voltaic efficiency term $\bar\eta_{V}$. A value of 0.842 was applied, as this is the voltaic efficiency at the mid-point of the permitted current-density range (\SI{1600}{\ampere\per\square\meter}). Similarly, the NLP formulation in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} requires values for area specific resistance ($ASR$) and Faradaic potential ($V_a$) in order to parametrise voltaic losses in the objective function. These were derived by multiplying $OCV_{50\%}$ by $1-\sqrt{\eta_V}$ to obtain an over-potential $\delta_v$ at the three current-density points in the experimental data, assuming losses are symmetrical in charge and discharge. $ASR$ was obtained as the gradient of the linear regression on this plot, and $V_a$ as the intercept at zero current-density.

\cref{fig:Va_ASR_Derivation} shows the regression of $\delta_V$ against current density for the data taken from \cite{Reed2016}.

\begin{figure}[!ht]
\centering
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./ASR_deriv.pdf}
\caption{Derivation of $ASR$ (gradient) and $V_a$ (intercept) for IDD2s VRFB embodiment reported by \cite{Reed2016} at 400 \si{\ml\per\minute}.}
\label{fig:Va_ASR_Derivation}
\end{figure}

>>>>

\begin{document}
\linenumbers

\title{MILP for idling to minimise coulombic losses}

\section{MILP formulation with fixed coulombic losses}
In the NLP formulation introduced in XX, it is assumed that the coulombic efficiency of the RFB is constant, i.e. the coulombic losses are proportional to the useful current density. \textit{What I say in assumptions section of paper...}

In order to address this concern, the SOC at time $t$ was redefined by:

\begin{equation}
\label{eqn: Method_MINLP_SOC_Tracker}
SOC_t = SOC_{t-1} +\frac{A.\tau}{1000C}(I_{C, t} - I_{D, t} - \delta_tl_C) \forall t \in Y
\end{equation}

Where $\delta_t$ is a binary indicator variable with 0 indicating an idle state and 1 an active state, and $l_C$ the fixed coulombic loss term (\si{\ampere}). It is assumed that the volume of electrolyte in the stack is negligible compared to the volume in the reservoirs.

The condition that the charging or discharging of the RFB requires the system to be in the active state is formalised by:

\begin{equation}
\begin{split}
    \label{eqn: On_off_condition}
    I_{C,t} > 0 \implies \delta_t = 1\\
    I_{D,t} > 0 \implies \delta_t = 1
    \end{split}
\end{equation}

The constraints that enforce these conditions are defined by:

\begin{equation}
\begin{split}
    \label{eqn: on_off_constraint}
    I_{C,t} - I_{Max}\delta_{C,t} \leq 0\\ 
    I_{D,t} - I_{Max}\delta_{D,t} \leq 0
    \end{split}
\end{equation}

 after \cite{Williams2013}.

 The expression of coulombic losses as a fixed term results in the restriction of operation in the low current-density range, where the coulombic losses would cancel out the voltaic loss reductions. \cref{fig:MINLP_on_EffC} shows a comparison of the optimal schedules obtained using the NLP formulation in XX and the MINLP formulation.

 \begin{figure}[!ht]
\centering
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/MINLP_schedule.pdf}
\caption{Optimal schedules obtained by NLP and MINLP optimisation for N2EX electrical price on 22$^{nd}$ August 2017.}
\label{fig:MINLP_on_EffC}
\end{figure}

A similar approach may be employed to deal with other fixed losses, such as the pumping power loss for a constant output pump.
\end{document}
